Distributional Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain

### 2018-11-03 pkg/disttree/inst/draft/draft_AOAS.tex
draft in AOAS format
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2377 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 4a0dc4a9a1926a0cd376361f2b456085878127c4

\documentclass[nojss]{jss}
\shortcites{Hamill+Bates+Whitaker:2013}

@@

%% packages
\usepackage{amstext,amsfonts,amsmath,bm,thumbpdf,lmodern,hyperref}
\usepackage[all]{hypcap}

%% tikz
\usepackage{array,makecell,tikz,color,soul}
\usetikzlibrary{arrows.meta,positioning,shapes,arrows,decorations.pathreplacing,calc,automata,mindmap}

%% need no \usepackage{Sweave}
\SweaveOpts{engine = R, concordance = FALSE, eps = FALSE, keep.source = TRUE, eval = TRUE}

%% commands
\newcommand{\fixme}[1]{\emph{\marginpar{FIXME} (#1)}}
\renewcommand{\Prob}{\mathbb{P} }
\renewcommand{\E}{\mathbb{E}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\Var}{\mathbb{V}}
\newcommand{\R}{\mathbb{R} }
\newcommand{\N}{\mathbb{N} }
\newcommand{\C}{\mathbb{C} }
\newcommand{\argmin}{\operatorname{argmin}\displaylimits}
\newcommand{\argmax}{\operatorname{argmax}\displaylimits}
\newcommand{\LS}{\mathcal{L}_n}
\newcommand{\TS}{\mathcal{T}_n}
\newcommand{\LSc}{\mathcal{L}_{\text{comb},n}}
\newcommand{\LSbc}{\mathcal{L}^*_{\text{comb},n}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\A}{\mathcal{A}}
\newcommand{\yn}{y_{\text{new}}}
\newcommand{\z}{\mathbf{z}}
\newcommand{\X}{\mathbf{X}}
\newcommand{\Y}{\mathbf{Y}}
\newcommand{\sX}{\mathcal{X}}
\newcommand{\sY}{\mathcal{Y}}
\newcommand{\T}{\mathbf{T}}
\newcommand{\x}{\mathbf{x}}
\renewcommand{\a}{\mathbf{a}}
\newcommand{\xn}{\mathbf{x}_{\text{new}}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\ws}{\mathbf{w}_\cdot}
\renewcommand{\t}{\mathbf{t}}
\newcommand{\M}{\mathbf{M}}
\renewcommand{\vec}{\text{vec}}
\newcommand{\B}{\mathbf{B}}
\newcommand{\K}{\mathbf{K}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\I}{\mathbf{I}}
\newcommand{\bS}{\mathbf{S}}
\newcommand{\cellx}{\pi_n[\x]}
\newcommand{\partn}{\pi_n(\mathcal{L}_n)}
\newcommand{\err}{\text{Err}}
\newcommand{\ea}{\widehat{\text{Err}}^{(a)}}
\newcommand{\ecv}{\widehat{\text{Err}}^{(cv1)}}
\newcommand{\ecvten}{\widehat{\text{Err}}^{(cv10)}}
\newcommand{\eone}{\widehat{\text{Err}}^{(1)}}
\newcommand{\eplus}{\widehat{\text{Err}}^{(.632+)}}
\newcommand{\eoob}{\widehat{\text{Err}}^{(oob)}}
\newcommand{\bft}{\mathbf{t}}

\definecolor{changes}{rgb}{1,1,0.7}
\newcommand{\hlc}[1]{{\sethlcolor{changes}\hl{#1}}}

\numberwithin{equation}{subsection}

<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", useFancyQuotes = FALSE, width = 70)
library("disttree")
library("crch")
library("countreg")
library("gamlss")
library("gamlss.cens")
gen.cens(NO, type = "left")
library("RainTyrol")
library("scoringRules")

## HCL palette
pal <- hcl(c(10, 128, 260, 290, 50), 100, 50)
pallight <- hcl(c(10, 128, 260, 290, 70), 100, 50, alpha = 0.25)
transpgray <- rgb(0.190,0.190,0.190, alpha = 0.2)

# set function for parallelization
applyfun <- function(X, FUN, ...) parallel::mclapply(X, FUN, ..., mc.cores = pmax(1, parallel::detectCores() - 1))

@

\title{Distributional Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain}
\Shorttitle{Distributional Regression Forests for Probabilistic Precipitation Forecasting}

\author{Lisa Schlosser\\Universit\"at Innsbruck
   \And Torsten Hothorn\\Universit\"at Z\"urich
   \And Reto Stauffer\\Universit\"at Innsbruck
   \And Achim Zeileis\\Universit\"at Innsbruck}
\Plainauthor{Lisa Schlosser, Torsten Hothorn, Reto Stauffer, Achim Zeileis}

\Abstract{
To obtain a probabilistic model for a dependent variable based on some set of
explanatory variables, a distributional approach is often adopted where the
parameters of the distribution are linked to regressors. In many classical
models this only captures the location of the distribution but
over the last decade there has been increasing interest in distributional
regression approaches modeling all parameters including location, scale, and
shape. Notably, so-called non-homogeneous Gaussian regression (NGR) models both
mean and variance of a Gaussian response and is particularly
popular in weather forecasting. 
Moreover, generalized additive models for location, scale, and shape (GAMLSS)
provide a framework where each distribution parameter is modeled separately
capturing smooth linear or nonlinear effects.
%
However, when variable selection is required and/or there are
non-smooth dependencies or interactions (especially unknown or of high-order),
it is challenging to establish a good GAMLSS. A natural alternative in these
situations would be the application of regression trees or random forests but,
so far, no general distributional framework is available for these. Therefore,
a framework for distributional regression trees and forests is proposed that blends
regression trees and random forests with classical distributions from the GAMLSS
framework as well as their censored or truncated counterparts.
%
To illustrate these novel approaches in practice, they are employed to obtain
probabilistic precipitation forecasts at numerous sites in a mountainous
region (Tyrol, Austria) based on a large number of numerical weather
prediction quantities. It is shown that the novel distributional regression forests
automatically select variables and interactions, performing on par or often even
better than GAMLSS specified either through prior meteorological knowledge or a
computationally more demanding boosting approach.
}
\Keywords{parametric models, regression trees, random forests, recursive partitioning, probabilistic forecasting, GAMLSS}

\Address{
  Lisa Schlosser, Reto Stauffer, Achim Zeileis \\
  Universit\"at Innsbruck \\
  Department of Statistics \\
  Faculty of Economics and Statistics \\
  Universit\"atsstr.~15 \\
  6020 Innsbruck, Austria \\
  E-mail: \email{Lisa.Schlosser@uibk.ac.at}, \email{Reto.Stauffer@uibk.ac.at},\\
  \email{Achim.Zeileis@R-project.org} \\
  URL: \url{https://www.uibk.ac.at/statistics/personal/schlosser-lisa/},\\
  \phantom{URL: }\url{https://retostauffer.org/},\\
  \phantom{URL: }\url{https://eeecon.uibk.ac.at/~zeileis/}\\

  Torsten Hothorn\\
  Universit\"at Z\"urich \\
  Institut f\"ur Epidemiologie, Biostatistik und Pr\"avention \\
  Hirschengraben 84\\
  CH-8001 Z\"urich, Switzerland \\
  E-mail: \email{Torsten.Hothorn@R-project.org}\\
  URL: \url{http://user.math.uzh.ch/hothorn/}\\

}

@@

@@

\pagebreak

@@

\begin{figure}[t!]
\begin{tikzpicture}
\node (a) at (0,0){
\minipage{0.29\textwidth}
\centering
<<motivation_GLM, echo=FALSE, results=hide>>=
nobs <- 200
## GLM
set.seed(7)
x <- c(1:nobs)/nobs
ytrue <- 1+x*1.5
y <- ytrue + rnorm(nobs,0,0.3)
@

<<plot_motivation_GLM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal[1], lwd = 7, main = "")
@
\endminipage};
\node (b) at (5,0) 
{
\minipage{0.29\textwidth}
\centering
<<motivation_GAM, echo=FALSE, results=hide>>=
## GAM
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3 
y <- ytrue + rnorm(nobs,0,0.3)
@

<<plot_motivation_GAM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal[1], lwd = 7, main = "")
@
\endminipage};
\node (c) at (10,0) 
{
\minipage{0.29\textwidth}
\centering
<<motivation_GAMLSS, echo=FALSE, results=hide>>=
## GAMLSS
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3
var <- exp(-(2*x)^2)/2
y <- ytrue + rnorm(nobs, 0, 0.1 + var)
@

<<plot_motivation_GAMLSS, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x, y, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
polygon(c(x, rev(x)), c(ytrue + 0.1 + var, rev(ytrue - 0.1 - var)),
  col = pallight[1], border = "transparent")
lines(x, ytrue, col = pal[1], lwd=7)
points(x, y, col = "slategray", pch = 19)
box(lwd = 5)
@
\endminipage};
\node (d) at (0,-2) 
{
\minipage{0.29\textwidth}
\centering
\vspace{0.0cm}
LM, GLM  
\endminipage};
\node (e) at (5,-2) 
{
\minipage{0.29\textwidth}
\centering
\vspace{0.0cm}
GAM 
\endminipage};
\node at (10,-2)
{
\minipage{0.28\textwidth}
\centering
\vspace{0.0cm}
GAMLSS 
\endminipage};
\draw[-{>[scale=4, length=2, width=3]},line width=0.4pt](2.1,0)--(2.9,0);
\draw[-{>[scale=4, length=2, width=3]},line width=0.4pt](7.1,0)--(7.9,0);
\end{tikzpicture}
\caption{\label{fig:gam}Parametric modeling developments. (Generalized) linear models (left),
generalized additive models (middle), generalized additive models for location, scale, and shape (right).}
\end{figure}

@@

\begin{figure}[t!]
\begin{tikzpicture}
\node (a) at (0,0){
\minipage{0.29\textwidth}
\centering
<<motivation_regtree, echo=FALSE, results=hide>>=
## Reg. Tree
set.seed(7)
kappa <- 12
x <- c(1:nobs)/nobs
ytrue <- ytree <- yforest <- numeric(length = length(x))
for(i in 1:nobs) ytrue[i] <- if(x[i]<1/3) 0.5 else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))
y <- ytrue + rnorm(nobs,0,0.3)
for(i in 1:nobs) ytree[i] <- if(x[i]<1/3) 0.5 else {if(x[i]<2/3) 2 else 1}
@

<<plot_motivation_regtree, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
#lines(x = x, y = ytrue, col = "gray", lwd=5, main = "")
lines(x = x, y = ytree, col = pal[1], lwd=7)
@
\endminipage};
\node (b) at (5,0) 
{
\minipage{0.29\textwidth}
\centering
<<motivation_randforest, echo=FALSE, results=hide>>=
## Random Forest
for(i in 1:nobs) yforest[i] <- if(x[i]<0.27) 0.5 else { if(x[i]<0.39) 0.5 + 1.5*(plogis((x[i]-0.33)/6*700)) else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))}
@

<<plot_motivation_randforest, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
#lines(x = x, y = ytrue, col = "gray", lwd=5, main = "")
lines(x = x, y = yforest, col = pal[1], lwd=7, main = "")
@
\endminipage};
\node (c) at (10,0) 
{
\minipage{0.29\textwidth}
\centering
<<motivation_distforest, echo=FALSE, results=hide>>=
## distforest
set.seed(4723)
for(i in 1:nobs) yforest[i] <- if(x[i]<0.27) 0.5 else { if(x[i]<0.39) 0.5 + 1.5*(plogis((x[i]-0.33)/6*700)) else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))}
xt <- 2*(x-0.5)
var <- exp(-(2*xt)^4)/2
y <- ytrue + rnorm(nobs, 0, 0.1 + var)
x <- c(1:nobs)/nobs
@

<<plot_motivation_distforest, fig=TRUE, echo=FALSE>>=
par(mar = c(2, 0, 2, 0))
plot(x, y, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
usr <- par("usr")
rect(usr[1], usr[3], usr[2], usr[4], col = gray(0.94), border = "transparent")
polygon(c(x, rev(x)), c(ytrue + 0.1 + var, rev(ytrue - 0.1 - var)),
  col = pallight[1], border = "transparent")
lines(x = x, y = yforest, col = pal[1], lwd=7, main = "")
points(x, y, col = "slategray", pch = 19)
box(lwd=8)
@
\endminipage};
\node (d) at (0,-2) 
{
\minipage{0.29\textwidth}
\centering
\vspace{0.0cm}
Regression tree 
\endminipage};
\node (e) at (5,-2) 
{
\minipage{0.29\textwidth}
\centering
\vspace{0.0cm}
Random forest 
\endminipage};
\node at (10,-2)
{
\minipage{0.28\textwidth}
\centering
\vspace{0.0cm}
Distributional forest 
\endminipage};
\node (g) at (0,-5.4) 
{
\minipage{0.29\textwidth}
\vspace{0.2cm}
%\hspace{0.0cm}
\centering
<<motivation_disttree, echo=FALSE, results=hide>>=
## disttree
set.seed(74)
x <- c(1:nobs)/nobs
var <- ytree/3
y <- ytrue + rnorm(nobs,0,var)
@

<<plot_motivation_disttree, fig=TRUE, echo=FALSE>>=
par(mar = c(2, 0, 2, 0))
plot(x, y, xaxt="n", yaxt="n", ann=FALSE, type = "n")
usr <- par("usr")
rect(usr[1], usr[3], usr[2], usr[4], col = gray(0.94), border = "transparent")
lines(x = x, y = ytree, col = pal[1], lwd=7, main = "")
polygon(c(x, rev(x)), c(ytree + var, rev(ytree - var)),
  col = pallight[1], border = "transparent")
points(x, y, col = "slategray", pch = 19)
box(lwd=8)
@
\vspace{-0.2cm}%\\
Distributional tree
\vspace{0.2cm}
\endminipage};
\draw[-{>[scale=4, length=2, width=3]},line width=0.4pt](2.1,0)--(2.9,0);
\draw[-{>[scale=4, length=2, width=3]},line width=0.4pt](7.1,0)--(7.9,0);
\draw[-{>[scale=4, length=2, width=3]},line width=0.4pt](0,-2.4)--(0,-3.2);
\draw[-{>[scale=4, length=2, width=3]},line width=0.4pt](2.2,-5.4)--(7.9,-2.5);
\end{tikzpicture}
\caption{\label{fig:tree}Tree and forest developments.
Regression tree (top left), distributional tree (bottom left), random forest
(top middle), and distributional forest (top right).}
\end{figure}

In summary, the models discussed so far provide a broad and powerful toolset for
parametric distributional fits depending on a specified set of additive linear or
smooth nonlinear terms. A rather different approach to capturing the dependence on
covariates are tree-based models.
\begin{itemize}
  \item \emph{Regression trees} (\citealp{Breiman+Friedman+Stone:1984}) recursively
    split the data into more homogeneous subgroups and can thus capture abrupt shifts
    (Figure~\ref{fig:tree}, top left) and approximate nonlinear functions.
    Furthermore, trees automatically carry out a forward selection of covariates and
    their interactions. 
  \item \emph{Random forests} (\citealp{Breiman:2001}) average the predictions of
    an ensemble of trees fitted to resampled versions of the learning data. This
    stabilizes the recursive partitions from individual trees and hence better
    approximates smooth functions (Figure~\ref{fig:tree}, top middle).
\end{itemize}
While classical regression trees and random forests only model the mean of the response
we propose to follow the ideas from GAMLSS modeling -- as outlined
in Figure~\ref{fig:gam} -- and combine tree-based methods with parametric distributional models,
yielding two novel techniques:
\begin{itemize}
  \item \emph{Distributional regression trees} (for short: \emph{distributional trees}) split the data into more homogeneous groups with
    respect to a parametric distribution, thus capturing changes in any distribution
    parameter like location, scale, or shape (Figure~\ref{fig:tree}, bottom left).
  \item \emph{Distributional regression forests} (for short: \emph{distributional forests}) utilize an ensemble of distributional trees
    for obtaining stabilized and smoothed parametric predictions (Figure~\ref{fig:tree},
    top right).
\end{itemize}

@@

<<echo=FALSE, results=hide>>=
#### Axams prediction 24 - 4
if(file.exists("Axams_24to4.rda")){
  load("Axams_24to4.rda")
} else {
  source("Axams_24to4.R")
  Axams_24to4 <- Axams_24to4()
  save(Axams_24to4, file = "Axams_24to4.rda")
}

#### prepare data for plot of estimated density functions
# predictions for one day (in each of the four years) 
# (19th of July 2011 is missing)
pday <- 24  # 2 (hohe Beobachtung zu niedrig geschaetzt), 4, 15, evtl. auch 7, 8, 23 (eine 0-Beobachtung und 2 sehr aehnliche),

pdays <- if(pday<19) c(pday, pday + 31, pday + 62, pday + 92) else c(pday, pday + 31, pday + 61, pday + 92)
pdf <- predict(Axams_24to4$df, newdata = Axams_24to4$testdata[pdays,], type = "parameter")
df_mu <- pdf$mu
df_sigma <- pdf$sigma
df_exp <- pnorm(df_mu/df_sigma) * (df_mu + df_sigma * (dnorm(df_mu/df_sigma) / pnorm(df_mu/df_sigma)))
cbind(df_exp, Axams_24to4$testdata[pdays,"robs"])

# plot predicted distributions together with observations
set.seed(7)
x <- c(0.01, sort(runif(500,0.01,8)))
y1 <- crch::dcnorm(x, mean = df_mu[1], sd = df_sigma[1], left = 0)
y2 <- crch::dcnorm(x, mean = df_mu[2], sd = df_sigma[2], left = 0)
y3 <- crch::dcnorm(x, mean = df_mu[3], sd = df_sigma[3], left = 0)
y4 <- crch::dcnorm(x, mean = df_mu[4], sd = df_sigma[4], left = 0)
dayending <- if(pday > 3) "th" else switch(pday, "1" = {dayending <- "st"}, "2" = {dayending <- "nd"}, "3" = {dayending <- "rd"})

# switch x-axis back to untransformed scale
x <- x^(1.6)

# point mass (slightly shifted)
pm1 <- c(0.15, crch::dcnorm(-1, mean = df_mu[1], sd = df_sigma[1], left = 0))     #  0.05
pm2 <- c(0.05, crch::dcnorm(-1, mean = df_mu[2], sd = df_sigma[2], left = 0))     #  0.01
pm3 <- c(-0.05, crch::dcnorm(-1, mean = df_mu[3], sd = df_sigma[3], left = 0))    # -0.03
pm4 <- c(-0.15, crch::dcnorm(-1, mean = df_mu[4], sd = df_sigma[4], left = 0))    # -0.07

# observations
obs1 <- c(Axams_24to4$testdata[pdays,"robs"][1]^(1.6), crch::dcnorm(Axams_24to4$testdata[pdays,"robs"][1], mean = df_mu[1], sd = df_sigma[1], left = 0))
obs2 <- c(Axams_24to4$testdata[pdays,"robs"][2]^(1.6), crch::dcnorm(Axams_24to4$testdata[pdays,"robs"][2], mean = df_mu[2], sd = df_sigma[2], left = 0))
obs3 <- c(Axams_24to4$testdata[pdays,"robs"][3]^(1.6), crch::dcnorm(Axams_24to4$testdata[pdays,"robs"][3], mean = df_mu[3], sd = df_sigma[3], left = 0))
obs4 <- c(Axams_24to4$testdata[pdays,"robs"][4]^(1.6), crch::dcnorm(Axams_24to4$testdata[pdays,"robs"][4], mean = df_mu[4], sd = df_sigma[4], left = 0))

#legendheight
lh1 <- crch::dcnorm(0.01, mean = df_mu[1], sd = df_sigma[1], left = 0)
lh2 <- crch::dcnorm(0.01, mean = df_mu[2], sd = df_sigma[2], left = 0)
lh3 <- crch::dcnorm(0.01, mean = df_mu[3], sd = df_sigma[3], left = 0)
lh4 <- crch::dcnorm(0.01, mean = df_mu[4], sd = df_sigma[4], left = 0)

############
# get predicted parameter of all models for testdata

# distributional tree
pdt <- predict(Axams_24to4$dt, newdata = Axams_24to4$testdata, type = "parameter")
dt_mu <- pdt$mu
dt_sigma <- pdt$sigma
dt_exp <- pnorm(dt_mu/dt_sigma) * (dt_mu + dt_sigma * (dnorm(dt_mu/dt_sigma) / pnorm(dt_mu/dt_sigma)))

# distributional forest
pdf <- predict(Axams_24to4$df, newdata = Axams_24to4$testdata, type = "parameter")
df_mu <- pdf$mu
df_sigma <- pdf$sigma
df_exp <- pnorm(df_mu/df_sigma) * (df_mu + df_sigma * (dnorm(df_mu/df_sigma) / pnorm(df_mu/df_sigma)))

# prespecified GAM
g_mu <- predict(Axams_24to4$g, newdata = Axams_24to4$testdata, what = "mu", type = "response", data = Axams_24to4$g_learndata)
g_sigma <- predict(Axams_24to4$g, newdata = Axams_24to4$testdata, what = "sigma", type = "response", data = Axams_24to4$g_learndata)
g_exp <- pnorm(g_mu/g_sigma) * (g_mu + g_sigma * (dnorm(g_mu/g_sigma) / pnorm(g_mu/g_sigma)))

# boosted GAM
pgb <- predict(Axams_24to4$gb, newdata = Axams_24to4$testdata, parameter = c("mu","sigma"), type = "response")
gb_mu <- pgb$mu
gb_sigma <- pgb$sigma
gb_exp <- pnorm(gb_mu/gb_sigma) * (gb_mu + gb_sigma * (dnorm(gb_mu/gb_sigma) / pnorm(gb_mu/gb_sigma)))

>>>>

% Template for the submission to:
%   The Annals of Probability           [aop]
%   The Annals of Applied Probability   [aap]
%   The Annals of Statistics            [aos] 
%   The Annals of Applied Statistics    [aoas]
%   Stochastic Systems                  [ssy]
%
%Author: In this template, the places where you need to add information
%        (or delete line) are indicated by {???}.  Mostly the information
%        required is obvious, but some explanations are given in lines starting
%Author:
%All other lines should be ignored.  After editing, there should be
%no instances of ??? after this line.

% use option [preprint] to remove info line at bottom
% journal options: aop,aap,aos,aoas,ssy
% natbib option: authoryear

\documentclass[aoas, preprint]{imsart}
%\shortcites{Hamill+Bates+Whitaker:2013}

@@

%% packages
\usepackage{amstext,amsfonts,amsmath,bm,thumbpdf,lmodern}
%\usepackage{amsthm,amsmath,natbib}
\usepackage{array,makecell,tikz}
\RequirePackage[OT1]{fontenc}
\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage[all]{hypcap}
\RequirePackage{natbib}

%% tikz
\usetikzlibrary{arrows.meta,positioning,shapes,arrows,decorations.pathreplacing,calc,automata,mindmap}
%\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}

%% colors
\definecolor{HighlightOrange}{rgb}{0.9490196,0.5725490,0.0000000}
\definecolor{HighlightBlue}{rgb}{0.4784314,0.7490196,0.9803922}

% provide arXiv number if available:
\arxiv{arXiv:1804.02921}

% put your definitions there:
\startlocaldefs
\newcommand{\argmax}{\operatorname{argmax}\displaylimits}
\newcommand{\argmin}{\operatorname{argmin}\displaylimits}
\newcommand{\fixme}[1]{\emph{\marginpar{FIXME} (#1)}}
\newcommand{\I}{\mathbf{I}}
\endlocaldefs
\numberwithin{equation}{subsection}

@@

\begin{frontmatter}

% "Title of the paper"
\title{Distributional Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain}
%\runtitle{Distributional Regression Forests for Probabilistic Precipitation Forecasting}
\runtitle{Distributional Regression Forests}

% indicate corresponding author with \corref{}
% \author{\fnms{John} \snm{Smith}\corref{}\ead[label=e1]{smith@foo.com}\thanksref{t1}}
% \thankstext{t1}{Thanks to somebody} 
% \address{line 1\\ line 2\\ printead{e1}}
% \affiliation{Some University}

\begin{aug}
\author{\fnms{Lisa} \snm{Schlosser}\thanksref{m1}\corref{}\ead[label=e1]{Lisa.Schlosser@uibk.ac.at}
\ead[label=u1,url]{https://www.uibk.ac.at/statistics/personal/schlosser-lisa/}},
\author{\fnms{Torsten} \snm{Hothorn}\thanksref{m2}\ead[label=e2]{Torsten.Hothorn@R-project.org}
\ead[label=u2,url]{http://user.math.uzh.ch/hothorn/}},
\author{\fnms{Reto} \snm{Stauffer}\thanksref{m1}\ead[label=e3]{Reto.Stauffer@uibk.ac.at}
\ead[label=u3,url]{https://retostauffer.org/}}
\and
\author{\fnms{Achim} \snm{Zeileis}\thanksref{m1}\ead[label=e4]{Achim.Zeileis@R-project.org}
\ead[label=u4,url]{https://eeecon.uibk.ac.at/\~{}zeileis/}}

\runauthor{L. Schlosser, T. Hothorn, R. Stauffer, A. Zeileis}

\affiliation{Universit\"at Innsbruck\thanksmark{m1} and Universit\"at Z\"urich\thanksmark{m2}}

\address{Universit\"at Innsbruck \\
  Department of Statistics \\
  Faculty of Economics and Statistics \\
  Universit\"atsstr.~15 \\
  6020 Innsbruck, Austria \\
\printead{e1}\\
\phantom{E-mail:\ }\printead*{e3}\\
\phantom{E-mail:\ }\printead*{e4}}
%\printead{u1}\\
%\phantom{URL:\ }\printead*{u3}\\
%\phantom{URL:\ }\printead*{u4}}

\address{Universit\"at Z\"urich \\
Institut f\"ur Epidemiologie,\\
Biostatistik und Pr\"avention \\
Hirschengraben 84\\
CH-8001 Z\"urich, Switzerland \\
\printead{e2}}
%\printead{u2}}
\end{aug}

\begin{abstract}

To obtain a probabilistic model for a dependent variable based on some set of
explanatory variables, a distributional approach is often adopted where the
parameters of the distribution are linked to regressors. In many classical
models this only captures the location of the distribution but
over the last decade there has been increasing interest in distributional
regression approaches modeling all parameters including location, scale, and
shape. Notably, so-called non-homogeneous Gaussian regression (NGR) models both
mean and variance of a Gaussian response and is particularly
popular in weather forecasting. 
Moreover, generalized additive models for location, scale, and shape (GAMLSS)
provide a framework where each distribution parameter is modeled separately
capturing smooth linear or nonlinear effects.
%
However, when variable selection is required and/or there are
non-smooth dependencies or interactions (especially unknown or of high-order),
it is challenging to establish a good GAMLSS. A natural alternative in these
situations would be the application of regression trees or random forests but,
so far, no general distributional framework is available for these. Therefore,
a framework for distributional regression trees and forests is proposed that blends
regression trees and random forests with classical distributions from the GAMLSS
framework as well as their censored or truncated counterparts.
%
To illustrate these novel approaches in practice, they are employed to obtain
probabilistic precipitation forecasts at numerous sites in a mountainous
region (Tyrol, Austria) based on a large number of numerical weather
prediction quantities. It is shown that the novel distributional regression forests
automatically select variables and interactions, performing on par or often even
better than GAMLSS specified either through prior meteorological knowledge or a
computationally more demanding boosting approach.
\end{abstract}

%parametric models, regression trees, random forests, recursive partitioning, probabilistic forecasting, GAMLSS

%\begin{keyword}[class=MSC]
%\kwd[Primary ]{60K35}
%\kwd{60K35}
%\kwd[; secondary ]{60K35}
%\end{keyword}

\begin{keyword}
\kwd{parametric models}
\kwd{regression trees}
\kwd{random forests}
\kwd{recursive partitioning}
\kwd{probabilistic forecasting}
\kwd{GAMLSS}
\end{keyword}

\end{frontmatter}

@@

@@

\begin{figure}[t!]
\begin{tikzpicture}
\node (a) at (0,0){
\minipage{0.29\textwidth}
\centering
\setkeys{Gin}{width=0.9\textwidth}
\includegraphics{draft-plot_motivation_GLM}
\endminipage};
\node (b) at (4.5,0) 
{
\minipage{0.29\textwidth}
\centering
\setkeys{Gin}{width=0.9\textwidth}
\includegraphics{draft-plot_motivation_GAM}
\endminipage};
\node (c) at (9,0) 
{
\minipage{0.29\textwidth}
\centering
\setkeys{Gin}{width=0.9\textwidth}
\includegraphics{draft-plot_motivation_GAMLSS}
\endminipage};
\node (d) at (0,-2) 
{
\minipage{0.29\textwidth}
\centering
\vspace{0.0cm}
LM, GLM  
\endminipage};
\node (e) at (4.5,-2) 
{
\minipage{0.29\textwidth}
\centering
\vspace{0.0cm}
GAM 
\endminipage};
\node at (9,-2)
{
\minipage{0.28\textwidth}
\centering
\vspace{0.0cm}
GAMLSS 
\endminipage};
\draw[-{>[scale=3, length=2, width=3]},line width=0.4pt](1.85,0)--(2.65,0);
\draw[-{>[scale=3, length=2, width=3]},line width=0.4pt](6.35,0)--(7.15,0);
\end{tikzpicture}
\caption{\label{fig:gam}Parametric modeling developments. (Generalized) linear models (left),
generalized additive models (middle), generalized additive models for location, scale, and shape (right).}
\end{figure}

@@

\begin{figure}[t!]
\begin{tikzpicture}
\node (a) at (0,0){
\minipage{0.29\textwidth}
\centering
\setkeys{Gin}{width=0.9\textwidth}
\includegraphics{draft-plot_motivation_regtree}
\endminipage};
\node (b) at (4.5,0) 
{
\minipage{0.29\textwidth}
\centering
\setkeys{Gin}{width=0.9\textwidth}
\includegraphics{draft-plot_motivation_randforest}
\endminipage};
\node (c) at (9,0) 
{
\minipage{0.29\textwidth}
\centering
\setkeys{Gin}{width=0.9\textwidth}
\includegraphics{draft-plot_motivation_distforest}
\endminipage};
\node (d) at (0,-2) 
{
\minipage{0.29\textwidth}
\centering
\vspace{0.0cm}
Regression tree 
\endminipage};
\node (e) at (4.5,-2) 
{
\minipage{0.29\textwidth}
\centering
\vspace{0.0cm}
Random forest 
\endminipage};
\node at (9,-2)
{
\minipage{0.28\textwidth}
\centering
\vspace{0.0cm}
Distributional forest 
\endminipage};
\node (g) at (0,-4.78) 
{
\minipage{0.29\textwidth}
\vspace{0.2cm}
%\hspace{0.0cm}
\centering
\setkeys{Gin}{width=0.9\textwidth}
\includegraphics{draft-plot_motivation_disttree}
\vspace{-0.2cm}%\\
Distributional tree
\vspace{0.2cm}
\endminipage};
\draw[-{>[scale=3, length=2, width=3]},line width=0.3pt](1.85,0)--(2.65,0);
\draw[-{>[scale=3, length=2, width=3]},line width=0.3pt](6.35,0)--(7.15,0);
\draw[-{>[scale=3, length=2, width=3]},line width=0.3pt](0,-2.35)--(0,-3);
\draw[-{>[scale=3, length=2, width=3]},line width=0.3pt](2.2,-4.78)--(7.15,-2.5);
\end{tikzpicture}
\caption{\label{fig:tree}Tree and forest developments.
Regression tree (top left), distributional tree (bottom left), random forest
(top middle), and distributional forest (top right).}
\end{figure}

In summary, the models discussed so far provide a broad and powerful toolset for
parametric distributional fits depending on a specified set of additive linear or
smooth nonlinear terms. A rather different approach to capturing the dependence on
covariates are tree-based models.
\begin{itemize}
  \item \emph{Regression trees} \citep{Breiman+Friedman+Stone:1984} recursively
    split the data into more homogeneous subgroups and can thus capture abrupt shifts
    (Figure~\ref{fig:tree}, top left) and approximate nonlinear functions.
    Furthermore, trees automatically carry out a forward selection of covariates and
    their interactions. 
  \item \emph{Random forests} \citep{Breiman:2001} average the predictions of
    an ensemble of trees fitted to resampled versions of the learning data. This
    stabilizes the recursive partitions from individual trees and hence better
    approximates smooth functions (Figure~\ref{fig:tree}, top middle).
\end{itemize}
While classical regression trees and random forests only model the mean of the response
we propose to follow the ideas from GAMLSS modeling -- as outlined
in Figure~\ref{fig:gam} -- and combine tree-based methods with parametric distributional models,
yielding two novel techniques:
\begin{itemize}
  \item \emph{Distributional regression trees} (for short: \emph{distributional trees}) split the data into more homogeneous groups with
    respect to a parametric distribution, thus capturing changes in any distribution
    parameter like location, scale, or shape (Figure~\ref{fig:tree}, bottom left).
  \item \emph{Distributional regression forests} (for short: \emph{distributional forests}) utilize an ensemble of distributional trees
    for obtaining stabilized and smoothed parametric predictions (Figure~\ref{fig:tree},
    top right).
\end{itemize}

@@

\begin{figure}[t!]
\centering
\setkeys{Gin}{width=0.8\textwidth}
\includegraphics{draft-rain_Axams_pred}
\caption{\label{fig:axams-dist}Total precipitation predictions by a 
distributional forest at station Axams for July~24 in 2009, 2010, 2011 and 
2012 learned on data from 1985--2008. Observations are non-negative and 
modeled by a Gaussian distribution left-censored at zero.
The observations are depicted by crosses and the predicted point mass
from the model by filled circles.}
\end{figure}
